---
title: "Functional Analysis"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

# Setup

```{r setup}

library(tidyverse)
library(clusterProfiler)
library(org.Hs.eg.db)
library(enrichplot)
library(pathview)
library(DOSE)

knitr::opts_chunk$set(fig.width = 10, dpi = 300)

dir.create(path = "./results/FA", showWarnings = FALSE)

```

# Results

## TLR10 light vs. dark 3h

### Gene ontology

```{r TLR3h_go, warning=FALSE}

out.dir <- "./results/FA/TLR10_light_3h_vs_TLR10_dark_3h/"

dir.create(out.dir)

res.TLR3h <- read.csv("results/DESeq/TLR10_light_3h_vs_TLR10_dark_3h_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.TLR3h <- res.TLR3h$log2FoldChange
names(gene_list.TLR3h) <- res.TLR3h$ENSG
  
set.seed(5892)
gseGO.TLR3h <- gseGO(gene_list.TLR3h,
                     OrgDb = org.Hs.eg.db,
                     keyType = "ENSEMBL",
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = TRUE)

gseGO.TLR3h <- setReadable(gseGO.TLR3h, "org.Hs.eg.db", "ENSEMBL")
gseGO.TLR3h

write.csv(data.frame(gseGO.TLR3h), file = paste0(out.dir, "gsea_go.csv"), row.names = F)

TLR3h.go.plot <- nrow(gseGO.TLR3h) > 0

```

```{r TLR3h_go_dotplot, eval=TLR3h.go.plot}

dotplot(gseGO.TLR3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLR3h_go_netplot, eval=TLR3h.go.plot, fig.height=10}

cnetplot(gseGO.TLR3h, showCategory = 5, color.params = list(foldChange = gene_list.TLR3h),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLR3h_go_mapplot, eval=TLR3h.go.plot, fig.height=9}
gseGO.TLR3h.pwts <- pairwise_termsim(gseGO.TLR3h)

emapplot(gseGO.TLR3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG (Kyoto Encyclopedia of Genes and Genomes)

```{r TLR3h_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.TLR3h), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.TLR3h.entrez <- gene_list.TLR3h[names(gene_list.TLR3h) %in% ids$ENSEMBL]
names(gene_list.TLR3h.entrez) <- ids$ENTREZID
gene_list.TLR3h.entrez <- gene_list.TLR3h.entrez[!duplicated(names(gene_list.TLR3h.entrez))]

```


```{r TLR3h_kegg, warning=FALSE}

set.seed(5923)
gseKEGG.TLR3h <- gseKEGG(gene_list.TLR3h.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         eps = 0,
                         maxGSSize = 500,
                         minGSSize = 15,
                         seed = T)

gseKEGG.TLR3h <- setReadable(gseKEGG.TLR3h, "org.Hs.eg.db", "ENTREZID")
gseKEGG.TLR3h

write.csv(data.frame(gseKEGG.TLR3h), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

TLR3h.kegg.plot <- nrow(gseKEGG.TLR3h) > 0

```


```{r TLR3h_kegg_dotplot, eval=TLR3h.kegg.plot, fid.width = 10, fig.height=10}

dotplot(gseKEGG.TLR3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLR3h_kegg_netplot, eval=TLR3h.kegg.plot, fig.height=10}

cnetplot(gseKEGG.TLR3h, showCategory = 5, color.params = list(foldChange = gene_list.TLR3h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLR3h_kegg_mapplot, eval=TLR3h.kegg.plot, fig.height=9}
gseKEGG.TLR3h.pwts <- pairwise_termsim(gseKEGG.TLR3h)

emapplot(gseKEGG.TLR3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLR3h_kegg_pathview, eval=TLR3h.kegg.plot}

pathview(gene.data = gene_list.TLR3h.entrez, pathway.id = gseKEGG.TLR3h@result$ID[1], species = "hsa")

filename <- paste0(gseKEGG.TLR3h@result$ID[1], ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r TLR3h_do, warning=FALSE}

set.seed(5534)
gseDO.TLR3h <- gseDO(gene_list.TLR3h.entrez,
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = T)

gseDO.TLR3h <- setReadable(gseDO.TLR3h, "org.Hs.eg.db", "ENTREZID")
gseDO.TLR3h

write.csv(data.frame(gseDO.TLR3h), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

TLR3h.DO.plot <- nrow(gseDO.TLR3h) > 0

```

```{r TLR3h_do_dotplot, eval=TLR3h.DO.plot}

dotplot(gseDO.TLR3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLR3h_do_netplot, eval=TLR3h.DO.plot}

cnetplot(gseDO.TLR3h, showCategory = 5, layout = "dh", color.params = list(foldChange = gene_list.TLR3h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLR3h_do_mapplot, eval=TLR3h.DO.plot}

gseDO.TLR3h.pwts <- pairwise_termsim(gseDO.TLR3h)

emapplot(gseDO.TLR3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

## TLR10 light vs. dark 16h

### Gene ontology

```{r TLR16h_go, warning=FALSE}

out.dir <- "./results/FA/TLR10_light_16h_vs_TLR10_dark_16h/"

dir.create(out.dir)

res.TLR16h <- read.csv("results/DESeq/TLR10_light_16h_vs_TLR10_dark_16h_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.TLR16h <- res.TLR16h$log2FoldChange
names(gene_list.TLR16h) <- res.TLR16h$ENSG
  
set.seed(4735)
gseGO.TLR16h <- gseGO(gene_list.TLR16h,
                     OrgDb = org.Hs.eg.db,
                     keyType = "ENSEMBL",
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = TRUE)

gseGO.TLR16h <- setReadable(gseGO.TLR16h, "org.Hs.eg.db", "ENSEMBL")
gseGO.TLR16h

write.csv(data.frame(gseGO.TLR16h), file = paste0(out.dir, "gsea_go.csv"), row.names = F)

TLR16h.go.plot <- nrow(gseGO.TLR16h) > 0

```

```{r TLR16h_go_dotplot, eval=TLR16h.go.plot, fig.width=10, fig.height=10}

dotplot(gseGO.TLR16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLR16h_go_netplot, eval=TLR16h.go.plot, fig.height=10}

cnetplot(gseGO.TLR16h, showCategory = 5, color.params = list(foldChange = gene_list.TLR16h),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLR16h_go_mapplot, eval=TLR16h.go.plot, fig.height=9}
gseGO.TLR16h.pwts <- pairwise_termsim(gseGO.TLR16h)

emapplot(gseGO.TLR16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG

```{r TLR16h_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.TLR16h), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.TLR16h.entrez <- gene_list.TLR16h[names(gene_list.TLR16h) %in% ids$ENSEMBL]
names(gene_list.TLR16h.entrez) <- ids$ENTREZID
gene_list.TLR16h.entrez <- gene_list.TLR16h.entrez[!duplicated(names(gene_list.TLR16h.entrez))]

```


```{r TLR16h_kegg, warning=FALSE}

set.seed(5923)
gseKEGG.TLR16h <- gseKEGG(gene_list.TLR16h.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         eps = 0,
                         maxGSSize = 500,
                         minGSSize = 15,
                         seed = T)

gseKEGG.TLR16h <- setReadable(gseKEGG.TLR16h, "org.Hs.eg.db", "ENTREZID")
gseKEGG.TLR16h

write.csv(data.frame(gseKEGG.TLR16h), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

TLR16h.kegg.plot <- nrow(gseKEGG.TLR16h) > 0

```

```{r TLR16h_kegg_dotplot, eval=TLR16h.kegg.plot, fid.width = 10, fig.height=10}

dotplot(gseKEGG.TLR16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLR16h_kegg_netplot, eval=TLR16h.kegg.plot, fig.height=10}

cnetplot(gseKEGG.TLR16h, showCategory = 5, color.params = list(foldChange = gene_list.TLR16h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLR16h_kegg_mapplot, eval=TLR16h.kegg.plot, fig.height=9}
gseKEGG.TLR16h.pwts <- pairwise_termsim(gseKEGG.TLR16h)

emapplot(gseKEGG.TLR16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLR16h_kegg_pathview, eval=TLR16h.kegg.plot}

pathview(gene.data = gene_list.TLR16h.entrez, pathway.id = gseKEGG.TLR16h@result$ID[1], species = "hsa")

filename <- paste0(gseKEGG.TLR16h@result$ID[1], ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r TLR16h_do, warning=FALSE}

set.seed(6246)
gseDO.TLR16h <- gseDO(gene_list.TLR16h.entrez,
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = T)

gseDO.TLR16h <- setReadable(gseDO.TLR16h, "org.Hs.eg.db", "ENTREZID")
gseDO.TLR16h

write.csv(data.frame(gseDO.TLR16h), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

TLR16h.DO.plot <- nrow(gseDO.TLR16h) > 0

```

```{r TLR16h_do_dotplot, eval=TLR16h.DO.plot}

dotplot(gseDO.TLR16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLR16h_do_netplot, eval=TLR16h.DO.plot}

cnetplot(gseDO.TLR16h, showCategory = 5, color.params = list(foldChange = gene_list.TLR16h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLR16h_do_mapplot, eval=TLR16h.DO.plot}

gseDO.TLR16h.pwts <- pairwise_termsim(gseDO.TLR16h)

emapplot(gseDO.TLR16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

## TLR10 light vs. wildtype light 3h

### Gene ontology

```{r TLRlightWT3h_go, warning=FALSE}

out.dir <- "./results/FA/TLR10_light_3h_vs_wildtype_light_3h/"

dir.create(out.dir)

res.TLRlightWT3h <- read.csv("results/DESeq/TLR10_light_3h_vs_wildtype_light_3h_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.TLRlightWT3h <- res.TLRlightWT3h$log2FoldChange
names(gene_list.TLRlightWT3h) <- res.TLRlightWT3h$ENSG
  
set.seed(4735)
gseGO.TLRlightWT3h <- gseGO(gene_list.TLRlightWT3h,
                     OrgDb = org.Hs.eg.db,
                     keyType = "ENSEMBL",
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = TRUE)

gseGO.TLRlightWT3h <- setReadable(gseGO.TLRlightWT3h, "org.Hs.eg.db", "ENSEMBL")
gseGO.TLRlightWT3h

write.csv(data.frame(gseGO.TLRlightWT3h), file = paste0(out.dir, "gsea_go.csv"), row.names = F)

TLRlightWT3h.go.plot <- nrow(gseGO.TLRlightWT3h) > 0

```

```{r TLRlightWT3h_go_dotplot, eval=TLRlightWT3h.go.plot}

dotplot(gseGO.TLRlightWT3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRlightWT3h_go_netplot, eval=TLRlightWT3h.go.plot, fig.height=10}

cnetplot(gseGO.TLRlightWT3h, showCategory = 5, color.params = list(foldChange = gene_list.TLRlightWT3h),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRlightWT3h_go_mapplot, eval=TLRlightWT3h.go.plot, fig.height=9}
gseGO.TLRlightWT3h.pwts <- pairwise_termsim(gseGO.TLRlightWT3h)

emapplot(gseGO.TLRlightWT3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG

```{r TLRlightWT3h_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.TLRlightWT3h), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.TLRlightWT3h.entrez <- gene_list.TLRlightWT3h[names(gene_list.TLRlightWT3h) %in% ids$ENSEMBL]
names(gene_list.TLRlightWT3h.entrez) <- ids$ENTREZID
gene_list.TLRlightWT3h.entrez <- gene_list.TLRlightWT3h.entrez[!duplicated(names(gene_list.TLRlightWT3h.entrez))]

```

 
```{r TLRlightWT3h_kegg, warning=FALSE}

set.seed(5923)
gseKEGG.TLRlightWT3h <- gseKEGG(gene_list.TLRlightWT3h.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         eps = 0,
                         maxGSSize = 500,
                         minGSSize = 15,
                         seed = T)

gseKEGG.TLRlightWT3h <- setReadable(gseKEGG.TLRlightWT3h, "org.Hs.eg.db", "ENTREZID")
gseKEGG.TLRlightWT3h

write.csv(data.frame(gseKEGG.TLRlightWT3h), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

TLRlightWT3h.kegg.plot <- nrow(gseKEGG.TLRlightWT3h) > 0

```

```{r TLRlightWT3h_kegg_dotplot, eval=TLRlightWT3h.kegg.plot, fid.width = 10, fig.height=10}

dotplot(gseKEGG.TLRlightWT3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRlightWT3h_kegg_netplot, eval=TLRlightWT3h.kegg.plot, fig.height=10}

cnetplot(gseKEGG.TLRlightWT3h, showCategory = 5, color.params = list(foldChange = gene_list.TLRlightWT3h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRlightWT3h_kegg_mapplot, eval=TLRlightWT3h.kegg.plot, fig.height=9}
gseKEGG.TLRlightWT3h.pwts <- pairwise_termsim(gseKEGG.TLRlightWT3h)

emapplot(gseKEGG.TLRlightWT3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRlightWT3h_kegg_pathview, eval=TLRlightWT3h.kegg.plot}

pathview(gene.data = gene_list.TLRlightWT3h.entrez, pathway.id = gseKEGG.TLRlightWT3h@result$ID[1], species = "hsa")

filename <- paste0(gseKEGG.TLRlightWT3h@result$ID[1], ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r TLRlightWT3h_do, warning=FALSE}

set.seed(6246)
gseDO.TLRlightWT3h <- gseDO(gene_list.TLRlightWT3h.entrez,
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = T)

gseDO.TLRlightWT3h <- setReadable(gseDO.TLRlightWT3h, "org.Hs.eg.db", "ENTREZID")
gseDO.TLRlightWT3h

write.csv(data.frame(gseDO.TLRlightWT3h), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

TLRlightWT3h.DO.plot <- nrow(gseDO.TLRlightWT3h) > 0

```

```{r TLRlightWT3h_do_dotplot, eval=TLRlightWT3h.DO.plot}

dotplot(gseDO.TLRlightWT3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRlightWT3h_do_netplot, eval=TLRlightWT3h.DO.plot}

cnetplot(gseDO.TLRlightWT3h, showCategory = 5, color.params = list(foldChange = gene_list.TLRlightWT3h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRlightWT3h_do_mapplot, eval=TLRlightWT3h.DO.plot}

gseDO.TLRlightWT3h.pwts <- pairwise_termsim(gseDO.TLRlightWT3h)

emapplot(gseDO.TLRlightWT3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

## TLR10 light vs. wildtype light 16h

### Gene ontology

```{r TLRlightWT16h_go, warning=FALSE}

out.dir <- "./results/FA/TLR10_light_16h_vs_wildtype_light_16h/"

dir.create(out.dir)

res.TLRlightWT16h <- read.csv("results/DESeq/TLR10_light_16h_vs_wildtype_light_16h_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.TLRlightWT16h <- res.TLRlightWT16h$log2FoldChange
names(gene_list.TLRlightWT16h) <- res.TLRlightWT16h$ENSG
  
set.seed(6548)
gseGO.TLRlightWT16h <- gseGO(gene_list.TLRlightWT16h,
                     OrgDb = org.Hs.eg.db,
                     keyType = "ENSEMBL",
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = TRUE)

gseGO.TLRlightWT16h <- setReadable(gseGO.TLRlightWT16h, "org.Hs.eg.db", "ENSEMBL")
gseGO.TLRlightWT16h

write.csv(data.frame(gseGO.TLRlightWT16h), file = paste0(out.dir, "gsea_go.csv"), row.names = F)

TLRlightWT16h.go.plot <- nrow(gseGO.TLRlightWT16h) > 0

```

```{r TLRlightWT16h_go_dotplot, eval=TLRlightWT16h.go.plot}

dotplot(gseGO.TLRlightWT16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRlightWT16h_go_netplot, eval=TLRlightWT16h.go.plot, fig.height=10}

cnetplot(gseGO.TLRlightWT16h, showCategory = 5, color.params = list(foldChange = gene_list.TLRlightWT16h),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRlightWT16h_go_mapplot, eval=TLRlightWT16h.go.plot, fig.height=9}
gseGO.TLRlightWT16h.pwts <- pairwise_termsim(gseGO.TLRlightWT16h)

emapplot(gseGO.TLRlightWT16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG

```{r TLRlightWT16h_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.TLRlightWT16h), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.TLRlightWT16h.entrez <- gene_list.TLRlightWT16h[names(gene_list.TLRlightWT16h) %in% ids$ENSEMBL]
names(gene_list.TLRlightWT16h.entrez) <- ids$ENTREZID
gene_list.TLRlightWT16h.entrez <- gene_list.TLRlightWT16h.entrez[!duplicated(names(gene_list.TLRlightWT16h.entrez))]

```


```{r TLRlightWT16h_kegg, warning=FALSE}

set.seed(5923)
gseKEGG.TLRlightWT16h <- gseKEGG(gene_list.TLRlightWT16h.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         eps = 0,
                         maxGSSize = 500,
                         minGSSize = 15,
                         seed = T)

gseKEGG.TLRlightWT16h <- setReadable(gseKEGG.TLRlightWT16h, "org.Hs.eg.db", "ENTREZID")
gseKEGG.TLRlightWT16h

write.csv(data.frame(gseKEGG.TLRlightWT16h), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

TLRlightWT16h.kegg.plot <- nrow(gseKEGG.TLRlightWT16h) > 0

```

```{r TLRlightWT16h_kegg_dotplot, eval=TLRlightWT16h.kegg.plot, fid.width = 10, fig.height=10}

dotplot(gseKEGG.TLRlightWT16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRlightWT16h_kegg_netplot, eval=TLRlightWT16h.kegg.plot, fig.height=10}

cnetplot(gseKEGG.TLRlightWT16h, showCategory = 5, color.params = list(foldChange = gene_list.TLRlightWT16h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRlightWT16h_kegg_mapplot, eval=TLRlightWT16h.kegg.plot, fig.height=9}
gseKEGG.TLRlightWT16h.pwts <- pairwise_termsim(gseKEGG.TLRlightWT16h)

emapplot(gseKEGG.TLRlightWT16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRlightWT16h_kegg_pathview, eval=TLRlightWT16h.kegg.plot}

pathview(gene.data = gene_list.TLRlightWT16h.entrez, pathway.id = gseKEGG.TLRlightWT16h@result$ID[1], species = "hsa")

filename <- paste0(gseKEGG.TLRlightWT16h@result$ID[1], ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r TLRlightWT16h_do, warning=FALSE}

set.seed(6246)
gseDO.TLRlightWT16h <- gseDO(gene_list.TLRlightWT16h.entrez,
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = T)

gseDO.TLRlightWT16h <- setReadable(gseDO.TLRlightWT16h, "org.Hs.eg.db", "ENTREZID")
gseDO.TLRlightWT16h

write.csv(data.frame(gseDO.TLRlightWT16h), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

TLRlightWT16h.DO.plot <- nrow(gseDO.TLRlightWT16h) > 0

```

```{r TLRlightWT16h_do_dotplot, eval=TLRlightWT16h.DO.plot}

dotplot(gseDO.TLRlightWT16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRlightWT16h_do_netplot, eval=TLRlightWT16h.DO.plot}

cnetplot(gseDO.TLRlightWT16h, showCategory = 5, color.params = list(foldChange = gene_list.TLRlightWT16h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRlightWT16h_do_mapplot, eval=TLRlightWT16h.DO.plot}

gseDO.TLRlightWT16h.pwts <- pairwise_termsim(gseDO.TLRlightWT16h)

emapplot(gseDO.TLRlightWT16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

## TLR10 dark vs. wildtype light 3h

### Gene ontology

```{r TLRdarkWT3h_go, warning=FALSE}

out.dir <- "./results/FA/TLR10_dark_3h_vs_wildtype_light_3h/"

dir.create(out.dir)

res.TLRdarkWT3h <- read.csv("results/DESeq/TLR10_dark_3h_vs_wildtype_light_3h_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.TLRdarkWT3h <- res.TLRdarkWT3h$log2FoldChange
names(gene_list.TLRdarkWT3h) <- res.TLRdarkWT3h$ENSG
  
set.seed(4735)
gseGO.TLRdarkWT3h <- gseGO(gene_list.TLRdarkWT3h,
                     OrgDb = org.Hs.eg.db,
                     keyType = "ENSEMBL",
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = TRUE)

gseGO.TLRdarkWT3h <- setReadable(gseGO.TLRdarkWT3h, "org.Hs.eg.db", "ENSEMBL")
gseGO.TLRdarkWT3h

write.csv(data.frame(gseGO.TLRdarkWT3h), file = paste0(out.dir, "gsea_go.csv"), row.names = F)

TLRdarkWT3h.go.plot <- nrow(gseGO.TLRdarkWT3h) > 0

```

```{r TLRdarkWT3h_go_dotplot, eval=TLRdarkWT3h.go.plot}

dotplot(gseGO.TLRdarkWT3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRdarkWT3h_go_netplot, eval=TLRdarkWT3h.go.plot, fig.height=10}

cnetplot(gseGO.TLRdarkWT3h, showCategory = 5, color.params = list(foldChange = gene_list.TLRdarkWT3h),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRdarkWT3h_go_mapplot, eval=TLRdarkWT3h.go.plot, fig.height=9}
gseGO.TLRdarkWT3h.pwts <- pairwise_termsim(gseGO.TLRdarkWT3h)

emapplot(gseGO.TLRdarkWT3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG

```{r TLRdarkWT3h_translate_ids, warning=FALSE}

ids <- bitr(names(gene_list.TLRdarkWT3h), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.TLRdarkWT3h.entrez <- gene_list.TLRdarkWT3h[names(gene_list.TLRdarkWT3h) %in% ids$ENSEMBL]
names(gene_list.TLRdarkWT3h.entrez) <- ids$ENTREZID
gene_list.TLRdarkWT3h.entrez <- gene_list.TLRdarkWT3h.entrez[!duplicated(names(gene_list.TLRdarkWT3h.entrez))]

```


```{r TLRdarkWT3h_kegg, warning=FALSE}

set.seed(5923)
gseKEGG.TLRdarkWT3h <- gseKEGG(gene_list.TLRdarkWT3h.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         eps = 0,
                         maxGSSize = 500,
                         minGSSize = 15,
                         seed = T)

gseKEGG.TLRdarkWT3h <- setReadable(gseKEGG.TLRdarkWT3h, "org.Hs.eg.db", "ENTREZID")
gseKEGG.TLRdarkWT3h

write.csv(data.frame(gseKEGG.TLRdarkWT3h), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

TLRdarkWT3h.kegg.plot <- nrow(gseKEGG.TLRdarkWT3h) > 0

```

```{r TLRdarkWT3h_kegg_dotplot, eval=TLRdarkWT3h.kegg.plot, fid.width = 10, fig.height=10}

dotplot(gseKEGG.TLRdarkWT3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRdarkWT3h_kegg_netplot, eval=TLRdarkWT3h.kegg.plot, fig.height=10}

cnetplot(gseKEGG.TLRdarkWT3h, showCategory = 5, color.params = list(foldChange = gene_list.TLRdarkWT3h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRdarkWT3h_kegg_mapplot, eval=TLRdarkWT3h.kegg.plot, fig.height=9}
gseKEGG.TLRdarkWT3h.pwts <- pairwise_termsim(gseKEGG.TLRdarkWT3h)

emapplot(gseKEGG.TLRdarkWT3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRdarkWT3h_kegg_pathview, eval=TLRdarkWT3h.kegg.plot}

pathview(gene.data = gene_list.TLRdarkWT3h.entrez, pathway.id = gseKEGG.TLRdarkWT3h@result$ID[1], species = "hsa")

filename <- paste0(gseKEGG.TLRdarkWT3h@result$ID[1], ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r TLRdarkWT3h_do, warning=FALSE}

set.seed(6246)
gseDO.TLRdarkWT3h <- gseDO(gene_list.TLRdarkWT3h.entrez,
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = T)

gseDO.TLRdarkWT3h <- setReadable(gseDO.TLRdarkWT3h, "org.Hs.eg.db", "ENTREZID")
gseDO.TLRdarkWT3h

write.csv(data.frame(gseDO.TLRdarkWT3h), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

TLRdarkWT3h.DO.plot <- nrow(gseDO.TLRdarkWT3h) > 0

```

```{r TLRdarkWT3h_do_dotplot, eval=TLRdarkWT3h.DO.plot}

dotplot(gseDO.TLRdarkWT3h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRdarkWT3h_do_netplot, eval=TLRdarkWT3h.DO.plot}

cnetplot(gseDO.TLRdarkWT3h, showCategory = 5, color.params = list(foldChange = gene_list.TLRdarkWT3h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRdarkWT3h_do_mapplot, eval=TLRdarkWT3h.DO.plot}

gseDO.TLRdarkWT3h.pwts <- pairwise_termsim(gseDO.TLRdarkWT3h)

emapplot(gseDO.TLRdarkWT3h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```


## TLR10 dark vs. wildtype light 16h

### Gene ontology

```{r TLRdarkWT16h_go, warning=FALSE}

out.dir <- "./results/FA/TLR10_dark_16h_vs_wildtype_light_16h/"

dir.create(out.dir)

res.TLRdarkWT16h <- read.csv("results/DESeq/TLR10_dark_16h_vs_wildtype_light_16h_unfiltered.csv") %>%
  as_tibble() %>%
  arrange(desc(log2FoldChange))

gene_list.TLRdarkWT16h <- res.TLRdarkWT16h$log2FoldChange
names(gene_list.TLRdarkWT16h) <- res.TLRdarkWT16h$ENSG
  
set.seed(4735)
gseGO.TLRdarkWT16h <- gseGO(gene_list.TLRdarkWT16h,
                     OrgDb = org.Hs.eg.db,
                     keyType = "ENSEMBL",
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = TRUE)

gseGO.TLRdarkWT16h <- setReadable(gseGO.TLRdarkWT16h, "org.Hs.eg.db", "ENSEMBL")
gseGO.TLRdarkWT16h

write.csv(data.frame(gseGO.TLRdarkWT16h), file = paste0(out.dir, "gsea_go.csv"), row.names = F)

TLRdarkWT16h.go.plot <- nrow(gseGO.TLRdarkWT16h) > 0

```

```{r TLRdarkWT16h_go_dotplot, eval=TLRdarkWT16h.go.plot}

dotplot(gseGO.TLRdarkWT16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRdarkWT16h_go_netplot, eval=TLRdarkWT16h.go.plot, fig.height=10}

cnetplot(gseGO.TLRdarkWT16h, showCategory = 5, color.params = list(foldChange = gene_list.TLRdarkWT16h),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRdarkWT16h_go_mapplot, eval=TLRdarkWT16h.go.plot, fig.height=9}
gseGO.TLRdarkWT16h.pwts <- pairwise_termsim(gseGO.TLRdarkWT16h)

emapplot(gseGO.TLRdarkWT16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")
```

### KEGG

```{r TLRdarkWT16h_translate_ids, warning=FALSE}
ids <- bitr(names(gene_list.TLRdarkWT16h), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
ids <- ids[!duplicated(ids$ENSEMBL),]

gene_list.TLRdarkWT16h.entrez <- gene_list.TLRdarkWT16h[names(gene_list.TLRdarkWT16h) %in% ids$ENSEMBL]
names(gene_list.TLRdarkWT16h.entrez) <- ids$ENTREZID
gene_list.TLRdarkWT16h.entrez <- gene_list.TLRdarkWT16h.entrez[!duplicated(names(gene_list.TLRdarkWT16h.entrez))]

```


```{r TLRdarkWT16h_kegg, warning=FALSE}

set.seed(5923)
gseKEGG.TLRdarkWT16h <- gseKEGG(gene_list.TLRdarkWT16h.entrez,
                         organism = "hsa",
                         keyType = "ncbi-geneid",
                         eps = 0,
                         maxGSSize = 500,
                         minGSSize = 15,
                         seed = T)

gseKEGG.TLRdarkWT16h <- setReadable(gseKEGG.TLRdarkWT16h, "org.Hs.eg.db", "ENTREZID")
gseKEGG.TLRdarkWT16h

write.csv(data.frame(gseKEGG.TLRdarkWT16h), file = paste0(out.dir, "gsea_kegg.csv"), row.names = F)

TLRdarkWT16h.kegg.plot <- nrow(gseKEGG.TLRdarkWT16h) > 0

```

```{r TLRdarkWT16h_kegg_dotplot, eval=TLRdarkWT16h.kegg.plot, fid.width = 10, fig.height=10}

dotplot(gseKEGG.TLRdarkWT16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRdarkWT16h_kegg_netplot, eval=TLRdarkWT16h.kegg.plot, fig.height=10}

cnetplot(gseKEGG.TLRdarkWT16h, showCategory = 5, color.params = list(foldChange = gene_list.TLRdarkWT16h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRdarkWT16h_kegg_mapplot, eval=TLRdarkWT16h.kegg.plot, fig.height=9}
gseKEGG.TLRdarkWT16h.pwts <- pairwise_termsim(gseKEGG.TLRdarkWT16h)

emapplot(gseKEGG.TLRdarkWT16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRdarkWT16h_kegg_pathview, eval=TLRdarkWT16h.kegg.plot}

pathview(gene.data = gene_list.TLRdarkWT16h.entrez, pathway.id = gseKEGG.TLRdarkWT16h@result$ID[1], species = "hsa")

filename <- paste0(gseKEGG.TLRdarkWT16h@result$ID[1], ".pathview.png")

file.rename(filename, paste0(out.dir, filename))

knitr::include_graphics(paste0(out.dir, filename))

```

### Disease ontology

```{r TLRdarkWT16h_do, warning=FALSE}

set.seed(6246)
gseDO.TLRdarkWT16h <- gseDO(gene_list.TLRdarkWT16h.entrez,
                     eps = 0,
                     maxGSSize = 500,
                     minGSSize = 15,
                     seed = T)

gseDO.TLRdarkWT16h <- setReadable(gseDO.TLRdarkWT16h, "org.Hs.eg.db", "ENTREZID")
gseDO.TLRdarkWT16h

write.csv(data.frame(gseDO.TLRdarkWT16h), file = paste0(out.dir, "gsea_do.csv"), row.names = F)

TLRdarkWT16h.DO.plot <- nrow(gseDO.TLRdarkWT16h) > 0

```

```{r TLRdarkWT16h_do_dotplot, eval=TLRdarkWT16h.DO.plot}

dotplot(gseDO.TLRdarkWT16h, showCategory = 15, size = NULL, split = ".sign", font.size = 9, label_format = 40) + 
  facet_grid(~.sign) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```

```{r TLRdarkWT16h_do_netplot, eval=TLRdarkWT16h.DO.plot}

cnetplot(gseDO.TLRdarkWT16h, showCategory = 5, color.params = list(foldChange = gene_list.TLRdarkWT16h.entrez),
         cex.params = list(category_label = 0.6, gene_label = 0.4)) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", limits = c(-2, 2), name = "Fold Change", oob = scales::squish)

```

```{r TLRdarkWT16h_do_mapplot, eval=TLRdarkWT16h.DO.plot}

gseDO.TLRdarkWT16h.pwts <- pairwise_termsim(gseDO.TLRdarkWT16h)

emapplot(gseDO.TLRdarkWT16h.pwts, force = 1.5,
         cex.params = list(category_label = 0.8)) +
  scale_fill_gradient(low = "red", high = "white", limits = c(0, 0.05), name = "adj. p-value")

```